---
title: "MCMC Practice Notebook"
author: "yangsong"
date: "10/25/2020"
output:
  html_document:
    toc: true
    theme: united
---


# Expectation of a truncated normal distribution
## Analytic Solution
- Do the integration of a standard normal pdf with range x > 3. The expected value is 0.00135

## Direct MC simulation
```{r, message=FALSE, warning=FALSE, fig.width=5, fig.height=5}
res <- c()
iter <- 10000

for (i in 1:iter){
  N <- 1000
  samples <- rnorm(N)
  res <- c(res, mean(samples > 3))
}

hist(res, col='skyblue3')
print(paste("means:", mean(res), sep=" "))
print(paste("variances:", var(res), sep=" "))
```

## Importance Sampling
```{r, message=FALSE, warning=FALSE, fig.width=5, fig.height=5}
res <- c()
iter <- 10000

for (i in 1:iter){
  N <- 1000
  samples <- rnorm(N, mean=4, sd=1)
  f <- dnorm(samples)
  g <- dnorm(samples, mean=4, sd=1)
  h <- samples > 3
  res <- c(res, mean(f*h/g))
}

hist(res, col='skyblue3')
print(paste("means:", mean(res), sep=" "))
print(paste("variances:", var(res), sep=" "))
```
- Better performance (low variance) compared with the direct monte carlo simulation



# Hierarchical Model
### Example Loading
```{r, message=FALSE, warning=FALSE, fig.width=5, fig.height=5}
#  This part of code is from here: http://avehtari.github.io/BDA_R_demos/demos_ch5/demo5_1.html
library(ggplot2)
library(gridExtra)
library(tidyr)

y <- c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,1,1,1,1,1,1,1,2,2,2,2,2,2,2,2,
        2,1,5,2,5,3,2,7,7,3,3,2,9,10,4,4,4,4,4,4,4,10,4,4,4,5,11,12,
        5,5,6,5,6,6,6,6,16,15,15,9,4)
n <- c(20,20,20,20,20,20,20,19,19,19,19,18,18,17,20,20,20,20,19,19,18,18,25,24,
       23,20,20,20,20,20,20,10,49,19,46,27,17,49,47,20,20,13,48,50,20,20,20,20,
       20,20,20,48,19,19,19,22,46,49,20,20,23,19,22,20,20,20,52,46,47,24,14)

x <- seq(0.0001, 0.9999, length.out = 1000)
bdens <- function(n, y, x){
    dbeta(x, y+1, n-y+1)  
}
```


### Separate model for each theta (theta assumes follow a beta distribution)
```{r, message=FALSE, warning=FALSE, fig.width=5, fig.height=5}
df_sep <- mapply(bdens, n, y, MoreArgs = list(x = x)) %>%
  as.data.frame() %>% cbind(x) %>% gather(ind, p, -x)

labs1 <- paste('posterior of', c('theta_j', 'theta_71'))
plot_sep <- ggplot(data = df_sep) +
  geom_line(aes(x = x, y = p, color = (ind=='V71'), group = ind)) +
  labs(x = expression(theta), y = '', title = 'Separate model', color = '') +
  scale_y_continuous(breaks = NULL) +
  scale_color_manual(values = c('blue','red'), labels = labs1) +
  theme(legend.background = element_blank(), legend.position = c(0.8,0.9))

plot_sep
```

### Pooled model
```{r, message=FALSE, warning=FALSE, fig.width=5, fig.height=5}
df_pool <- data.frame(x = x, p = dbeta(x, sum(y)+1, sum(n)-sum(y)+1))

plot_pool <- ggplot(data = df_pool) +
  geom_line(aes(x = x, y = p, color = '1')) +
  labs(x = expression(theta), y = '', title = 'Pooled model', color = '') +
  scale_y_continuous(breaks = NULL) +
  scale_color_manual(values = 'red', labels = 'Posterior of common theta') +
  theme(legend.background = element_blank(), legend.position = c(0.7,0.9))
plot_pool
```


### Marginal distribution of alpha and beta
```{r, message=FALSE, warning=FALSE, fig.width=5, fig.height=5}
A <- seq(1, 6, length.out = 100)
B <- seq(0, 30, length.out = 100)

# make grid point
cA <- rep(A, each = length(B))
cB <- rep(B, length(A))

#log_p(a,b)
lpfun <- function(a, b, y, n){
  log(a+b)*(-5/2) + sum(lgamma(a+b)-lgamma(a)-lgamma(b)+lgamma(a+y)+lgamma(b+n-y)-lgamma(a+b+n))
}
lp <- mapply(lpfun, cA, cB, MoreArgs = list(y, n))

#Subtract maximum value to avoid over/underflow in exponentation
#exp convert back to the exponential 
df_marg <- data.frame(x = cA, y = cB, p = exp(lp-max(lp)))


ggplot(data = df_marg, aes(x = x, y = y)) +
  geom_raster(aes(fill = p, alpha = p), interpolate = T) +
  geom_contour(aes(z = p), colour = 'black', size = 0.2) +
  coord_cartesian(xlim = c(1, 6), ylim = c(0, 30)) +
  ylab("alpha") +
  xlab("beta") + 
  scale_fill_gradient(low = 'yellow', high = 'red', guide = F) +
  scale_alpha(range = c(0, 1), guide = F)
```


### Sample from the grid




